---
output:
  md_document:
    variant: markdown_github
---

# Some convenient extensions to MeSH

MeSH (Medical Subject Headings) is a NLM controlled vocabulary thesaurus. The resource is structured as an ontology and is used for indexing articles for PubMed. Data sets included here provide more convenient access to this vocabulary as simple tables (in `rds` format).


## Datasets

### descriptor-terms

```{r}
readRDS('data/data_mesh_thesaurus.rds') |>
  head() |> knitr::kable()
```


### tree-numbers

```{r}
readRDS('data/data_mesh_trees.rds') |>
  head() |> knitr::kable()
```

### supplemental-terms

```{r}
readRDS('data/data_scr_thesaurus.rds') |>
  head() |> knitr::kable()
```


## Notes & useful links:

* [MeSH XML data elements](https://www.nlm.nih.gov/mesh/xml_data_elements.html)

* Supplementary Concept Records

> SCR records are created for some chemicals, drugs, and other concepts
such as rare diseases. They are labeled as "MeSH Supplementary Concept
Data" and the unique ID begins with the letter "C." 

> Supplementary Concept Records - these are not full MeSH Headings
and do not fall under the MeSH tree hierarchy. Many times they are used
to identify substances that are not included in the MeSH terms.

> These do not belong to the controlled vocabulary as such and are not used for indexing MEDLINE articles; instead they enlarge the thesaurus and contain links to the closest fitting descriptor to be used in a MEDLINE search. Many of these records describe chemical substances.


