---
output:
  md_document:
    variant: markdown_github
---


# MeSH resources (2026 update)

> **May 2026 update:** this repo now maintains only the core tables used by `puremoe`: `Descriptor Terms`, `Supplementary Concept Terms`, and `MeSH Tree Structures`.

> The R code for XML extraction and restructuring processes is available [here](https://github.com/jaytimm/mesh-builds/blob/main/descriptor-records-trees.Rmd). These data are utilized in the R package [puremoe](https://github.com/jaytimm/puremoe).



## MeSH ontology 

> Based on MeSH files `desc2026`, `mtrees2026` & `supp2026`.


### `descriptor-terms`

[Raw data](https://nlmpubs.nlm.nih.gov/projects/mesh/MESH_FILES/xmlmesh/)

```{r}
readRDS('data/data_mesh_thesaurus.rds') |>
  head() |> knitr::kable()
```


### `descriptor-tree-numbers`

[Raw data](https://nlmpubs.nlm.nih.gov/projects/mesh/MESH_FILES/meshtrees/)

```{r}
readRDS('data/data_mesh_trees.rds') |>
  head() |> knitr::kable()
```


### `supplemental-terms`

[Raw data](https://nlmpubs.nlm.nih.gov/projects/mesh/MESH_FILES/xmlmesh/)

```{r}
readRDS('data/data_scr_thesaurus.rds') |>
  head() |> knitr::kable()
```

## Maintained outputs

```{r eval=FALSE}
list.files('data', pattern = 'data_.*\\.rds$')
# [1] "data_mesh_thesaurus.rds" "data_mesh_trees.rds" "data_scr_thesaurus.rds"
```


### Notes & useful links:

* [MeSH XML data elements](https://www.nlm.nih.gov/mesh/xml_data_elements.html)

* XML elements as [RDF](https://id.nlm.nih.gov/mesh/D000001.html)

* [Utility functions](https://github.com/scienceai/mesh-tree)

* Supplementary Concept Records

> 'SCR records are created for some chemicals, drugs, and other concepts
such as rare diseases. They are labeled as "MeSH Supplementary Concept
Data" and the unique ID begins with the letter "C."' 

> 'Supplementary Concept Records - these are not full MeSH Headings
and do not fall under the MeSH tree hierarchy. Many times they are used
to identify substances that are not included in the MeSH terms.'

> 'These do not belong to the controlled vocabulary as such and are not used for indexing MEDLINE articles; instead they enlarge the thesaurus and contain links to the closest fitting descriptor to be used in a MEDLINE search. Many of these records describe chemical substances.'

